Approximation as Differentiation

### 2025-04-16 appendix/notes/biproduct.tex
Standardise on X,Y,Z,W for objects.

\begin{definition}[Biproduct]
Suppose $\cat{C}$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is
an object $A \biprod B$ of $C$ together with morphisms

\begin{center}
\begin{tikzcd}
   A \arrow[r, "\biinj_A", shift left] &
   A \biprod B \arrow[l, "\biproj_A", shift left] \arrow[r, "\biproj_B"', shift right] &
   B \arrow[l, "\biinj_B"', shift right]
\end{tikzcd}
\end{center}

@@

\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \biproj_A \comp \biinj_A &= \id_A \\
   \biproj_B \comp \biinj_A &= \zero_{A,B}
\end{salign*}
\end{center}
\end{minipage}%
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \biproj_B \comp \biinj_B &= \id_B \\
   \biproj_A \comp \biinj_B &= \zero_{B,A}
\end{salign*}
\end{center}
\end{minipage}

@@

\begin{salign*}
\biinj_A \comp \biproj_A + \biinj_B \comp \biproj_B &= \id_{A \biprod B}
\end{salign*}
\end{definition}

@@

We say a $\CMon$-enriched $\cat{C}$ is \emph{semi-additive} if it also has finite biproducts; having products
or coproducts is sufficient.

\begin{proposition}
\label{prop:biproduct:from-product-or-coproduct}
Suppose $\cat{C}$ is $\CMon$-enriched.
\begin{itemize}
\item If $\cat{C}$ has biproducts $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
\biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
\item If $\cat{C}$ has binary products $(A \times B, \proj_1, \proj_2)$ then:
\begin{enumerate}
\item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1}{f_2} + \prodM{g_1}{g_2}$
\item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
biproduct with $\biinj_A: A \to A \times B = \prodM{\id_A}{0_{A,B}}$ and $\biinj_B: B \to A \times B =
\prodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $\cat{C}$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
\begin{enumerate}
\item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1}{f_2} + \coprodM{g_1}{g_2}$
\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
\coprodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $\cat{C}$ has a terminal (or initial) object then it is a zero object.
\end{itemize}
\end{proposition}

>>>>

\begin{definition}[Biproduct]
Suppose $\cat{C}$ a $\CMon$-enriched category. For any objects $X, Y$, the \emph{biproduct} of $X$ and $Y$ is
an object $X \biprod Y$ of $C$ together with morphisms

\begin{center}
\begin{tikzcd}
   X \arrow[r, "\biinj_X", shift left] &
   X \biprod Y \arrow[l, "\biproj_X", shift left] \arrow[r, "\biproj_Y"', shift right] &
   Y \arrow[l, "\biinj_Y"', shift right]
\end{tikzcd}
\end{center}

@@

\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \biproj_X \comp \biinj_X &= \id_X \\
   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
\end{salign*}
\end{center}
\end{minipage}%
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \biproj_Y \comp \biinj_Y &= \id_Y \\
   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
\end{salign*}
\end{center}
\end{minipage}

@@

\begin{salign*}
(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
\end{salign*}
\end{definition}

@@

We call a $\CMon$-enriched $\cat{C}$ \emph{semi-additive} if it also has finite biproducts; having products or
coproducts is sufficient.

\begin{proposition}
\label{prop:biproduct:from-product-or-coproduct}
Suppose $\cat{C}$ is $\CMon$-enriched.
\begin{itemize}
\item If $\cat{C}$ has biproducts $(X \biprod Y, \biinj_X, \biinj_Y, \biproj_X, \biproj_Y)$ then $(X
\biprod Y, \biproj_X, \biproj_Y)$ is a product and $(X \biprod Y, \biinj_X, \biinj_Y)$ is a coproduct.
\item If $\cat{C}$ has binary products $(X \times Y, \proj_1, \proj_2)$ then:
\begin{enumerate}
\item $\prodM{0_{X,Y}}{0_{X,Z}} = 0_{X,Y \times Z}$
\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1}{f_2} + \prodM{g_1}{g_2}$
\item $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$ is a biproduct with $\biinj_X =
\prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y = \prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
\end{enumerate}
\item If $\cat{C}$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
\begin{enumerate}
\item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1}{f_2} + \coprodM{g_1}{g_2}$
\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
\coprodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $\cat{C}$ has a terminal (or initial) object then it is a zero object.
\end{itemize}
\end{proposition}

### 2025-04-16 appendix/notes/cmon-enriched.tex
Standardise on X,Y,Z,W for objects.

>>>>

Suppose $F, G: \cat{D} \to \cat{C}$ are functors where $\cat{C}$ is $\CMon$-enriched. Then the hom-object
$\Hom{\Func{\cat{C}}{\cat{D}}}{F}{G}$ of natural transformations between $F$ and $G$ has a zero $0_{F,G}$ and
addition $\eta + \mu$ given component-wise as $(0_{F,G})_X = 0_{F(X),G(X)}$ and $(\eta + \mu)_X = \eta_X +
\mu_X$.
